### 2021-03-22 ch07_conclusion/conclusion.tex
Cleaned abstract, introduction, scientific background and conclusion

The analysis of future climate projections from numerical climate model
simulations is of paramount importance for an accurate quantification of the
future climate change. Since this involves metrics of public interest like the
allowable fossil fuel emissions to meet particular warming targets, for example
the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015},
research in this field of science is not only relevant to climate scientists
but to policymakers and the whole human society. In the light of the large
spread of climate sensitivity in the most recent generation of climate models
from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
refinement of the output of multi-model climate projections is as relevant as
ever. This thesis quantifies the associated uncertainties, presents the
evaluation of established methods to reduce these uncertainties in the new
climate model ensemble and finally introduces a novel alternative technique
based on supervised \ac{ML}.

To answer the \namecrefs{enum:01:question_1} posed in
\cref{sec:01:key_science_questions} and to ensure a consistent evaluation of
the participating climate model ensembles in this thesis, the \ac{ESMValTool}
is used, an open-source community diagnostics and performance metrics tool for
the routine evaluation of \acp{ESM}. All analyses shown in this thesis are
performed with the \ac{ESMValTool}. Apart from that, further substantial
changes and additions to the code base of the tool have been implemented (see
\cref{ch:03:esmvaltool}), which lead to co-authorship in the scientific
documentation of the \ac{ESMValTool} \autocite{Eyring2020, Lauer2020,
  Righi2020, Weigel2020}. Since the open-source tool is freely available for
everyone, these improvements are beneficial for the entire scientific
community.

In a first analysis, the climate sensitivity metrics \ac{ECS} and \ac{TCR} are
evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved simulation of the shortwave \ac{CRE} in these
models when compared to observations. However, this change also substantially
reduces the strong negative shortwave cloud feedback over the Southern Ocean
that is present in previous \ac{CMIP} generations and that results from a cloud
phase change from ice clouds in the present-day to liquid clouds in the future.
In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
reduced by allowing for supercooled cloud liquid. As a consequence, these
models simulate less cloud ice over the present-day Southern Ocean than their
predecessor versions with no supercooled cloud liquid.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, suitable emergent constraints on the \acs{CMIP}6 ensemble are
evaluated. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. In total eleven emergent constraints on \ac{ECS} are
assessed, which are mostly related to cloud feedbacks since these constitute
the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
  Flato2013}. Since all of the evaluated emergent constraints have been derived
on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, out-of-sample tests on the emergent
constraints can be performed by assessing whether they still hold for the
\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
already published in \textcite{Schlund2020a}, a substantial reduction of skill
for the majority of emergent constraints is found when applied to the
\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
skill is expressed as a decrease of the coefficient of determination $R^2$ of
the emergent relationship and a decrease of the statistical significance using
the null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
confidence) for almost all emergent constraints, resulting in values of
$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
be answered with a \enquote{not very well} for the \acs{CMIP}6 ensemble. The
increased best estimates and spreads resulting from the emergent constraints in
\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
a published emergent constraint by \textcite{Wenzel2016}. This first step
corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2}, which shows a
range of $\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In
the second step, a \ac{ML}-based climate model weighting approach is used to
further constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. A a relationship between process‐oriented predictors
and future projections of \ac{GPP} is established and then utilized to project
today's observed conditions into the future. The prediction phase of the new
method can be interpreted as an implicit performance weighting. However, due to
the complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not
possible to extract specific values for the individual weights. Two target
variables are considered: the gridded monthly climatologies of absolute
\ac{GPP} (\range{2091}{2100}) and the gridded fractional \ac{GPP} change over
the \nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
answered with \enquote{yes} based on the results found.

@@

Climate change is a global phenomenon affecting the entire human society. Due
to the vastly complex nature of the Earth system, accurately quantifying future
climate change is a highly non-trivial task. This thesis covers several
important aspects of the climate system and metrics describing the changing
climate, but it is by no means complete and can be extended in an arbitrary
number of ways. This final \namecref{sec:07:outlook} presents several
suggestions on possible ways forward.

Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
strength of climate change. The range of model results for this important
climate metric, however, remains large and has not been narrowed down over the
last decades. In the latest generation of climate models contributing to
\acs{CMIP}6, the range has even increased.
\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
changes in the aerosol-cloud interaction and changes in the shortwave cloud
phase change feedback over the Southern Ocean. While these are likely
explanations, there might be additional relevant aspects that have not been
analyzed in detail yet. This could include additional feedback processes that
are not present in older \ac{CMIP} generations or missing feedback processes
that appear in previous \ac{CMIP} models but are absent in the \acs{CMIP}6
ensemble. Obvious candidates for these feedback processes are cloud-related
feedbacks, which pose a major source of uncertainty in climate sensitivity in
modern-day \acp{ESM} \autocite{Boucher2013}. Identifying and quantifying such
feedback mechanisms can further help to gain a better understanding of the
\acs{CMIP}6 models and potentially to reduce associated uncertainties in the
entire multi-model ensemble.

@@

The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
ever increasing computational power and resources provided by modern
supercomputers has the potential for enormous improvements in climate modeling
and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
weighting approach, further possible applications of \ac{AI} and \ac{ML} in
climate science involve for example \ac{ML}-based parametrizations that are
learned from high-resolution climate models \autocite{Rasp2018}, \ac{ML}-based
analysis and prediction of forcing patterns \autocite{Barnes2019,
  Mansfield2020} or learning of entire \acp{ESM} from observational products
\autocite{Geer2021}. Therefore, the foundation for innovative and
groundbreaking research in climate science for the near future has been laid,
which will hopefully help us in further understanding and fighting one of the
greatest challenges for humankind today: climate change.

>>>>

The analysis of future climate projections from numerical climate model
simulations is of paramount importance to assess future climate change under
different forcing scenarios. Since this involves metrics of public interest
like the allowable fossil fuel emissions to meet particular warming targets,
for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
\autocite{UNFCCC2015}, research in this field of science is not only relevant
to climate scientists but to policymakers and the whole human society. In the
light of the large spread of climate sensitivity in the most recent generation
of climate models from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical
evaluation and refinement of the output of multi-model climate projections is
as relevant as ever. This thesis quantifies the associated uncertainties,
presents the evaluation of established methods to reduce these uncertainties in
the new climate model ensemble and finally introduces a novel alternative
technique based on supervised \ac{ML}.

To answer the \namecrefs{enum:01:question_1} posed in
\cref{sec:01:key_science_questions} and to ensure a consistent evaluation of
the participating climate model ensembles in this thesis, the \ac{ESMValTool}
is used, an open-source community diagnostics and performance metrics tool for
the routine evaluation of \acp{ESM}. All analyses shown in this thesis are
performed with the \ac{ESMValTool}. Apart from that, further substantial
changes and additions to the code base of the tool have been implemented as
part of this thesis (see \cref{ch:03:esmvaltool}), which led to co-authorship
in the technical and scientific documentation of the \ac{ESMValTool}
\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}. Since the open-source
tool is freely available, the code that has been implemented as part of this
thesis is beneficial for the entire scientific community.

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. This change
leads to an improved simulation of the shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change due to warming is
reduced since these models simulate less cloud ice over the present-day
Southern Ocean than their predecessor versions with no supercooled cloud
liquid.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, already-published emergent constraints that have been derived
on models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
are evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints
use a physically-based inter-model relationship between an observable quantity
of the Earth system and a target variable to reduce uncertainties in the target
variable with observations \autocite{Allen2002}. In total eleven emergent
constraints on \ac{ECS} are assessed, which are mostly related to cloud
feedbacks since these constitute the most important source of uncertainty for
\ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated emergent
constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble,
out-of-sample tests on the emergent constraints can be performed by assessing
whether they still hold for the \acs{CMIP}6 models. In this study, which is
shown in \cref{ch:05:paper_ecs} and already published in
\textcite{Schlund2020a}, a substantial reduction of skill for the majority of
emergent constraints is found when applied to the \acs{CMIP}6 ensemble in
comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
decrease of the coefficient of determination $R^2$ of the emergent relationship
and a decrease of the statistical significance using the null hypothesis that
there is no correlation between the predictor and \ac{ECS}. Moreover, the
corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
emergent constraints. Thus, \cref{enum:01:question_2}
(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
\enquote{not very well} for the \acs{CMIP}6 ensemble. The increased best
estimates and spreads resulting from the emergent constraints in \acs{CMIP}6
are likely related to the increased \ac{MMM} and multi-model spread of \ac{ECS}
in \acs{CMIP}6. A possible reason for the reduced skill of the emergent
constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
a published emergent constraint by \textcite{Wenzel2016}. This first step
corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2}, which shows a
range of $\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In
the second step, a \ac{ML}-based climate model weighting approach is used to
further constrain the gridded \ac{GPP} based on present-day predictors that are
relevant for the simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach
is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
(gridded) target variables and non-linear relationships between the predictors
and the target variable. A a relationship between process‐oriented predictors
and future projections of \ac{GPP} is established and then utilized to project
today's observed conditions into the future. The prediction phase of the new
method can be interpreted as an implicit performance weighting. However, due to
the complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not
possible to extract specific values for the individual weights. Two target
variables are considered: the gridded monthly climatologies of absolute
\ac{GPP} (\range{2091}{2100}) and the gridded fractional \ac{GPP} change over
the \nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the dominant features. This study directly addresses
\cref{enum:01:question_3} (\enquote{\emph{\KeyScienceQuestionThree{}}}), which
can be confidently answered with \enquote{yes} based on the results found.

@@

Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
strength of climate change. The range of model results for this important
climate metric, however, remains large and has not been narrowed over the last
decades. In the latest generation of climate models contributing to
\acs{CMIP}6, the range has even increased \autocite{Meehl2020}.
\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
changes in the aerosol-cloud interaction and changes in the shortwave cloud
phase change feedback over the Southern Ocean. While these are likely
explanations, there might be additional relevant aspects that have not been
analyzed in detail yet. This could include additional feedback processes that
are not present in older \ac{CMIP} generations. Obvious candidates for these
feedback processes are cloud-related feedbacks, which pose a major source of
uncertainty in climate sensitivity in modern-day \acp{ESM}
\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
can further help to gain a better understanding of the \acs{CMIP}6 models and
potentially to reduce associated uncertainties in the entire multi-model
ensemble. A potentially interesting process relevant in this context is the
influence of the midlatitude jet position on clouds and \acp{CRE}
\autocite{Grise2016}.

@@

The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
ever increasing computational power and resources provided by modern
supercomputers has the potential for enormous improvements in climate modeling
and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
weighting approach, further possible applications of \ac{AI} and \ac{ML} in
climate science involve for example \ac{ML}-based parametrizations that are
learned from high-resolution climate models \autocite{Gentine2018, Rasp2018},
\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
  Mansfield2020} or learning of entire \acp{ESM} from observational products
\autocite{Geer2021}. Therefore, the foundation for innovative and
groundbreaking research in climate science for the near future has been laid,
which will help to further understand the Earth system and fight one of the
greatest challenges for humankind today: climate change.
